---
title: "Using checkpoint in a markdown document"
author: "Andrie de Vries"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using checkpoint in a markdown document}
  %\VignetteEngine{knitr::rmarkdown}
  %\usepackage[utf8]{inputenc}
---

# Introduction

Sometimes you want to use create a report using `markdown`, and you want to `checkpoint` the code in this document.

However, running this as part of a `RMarkdown` process is problematic, since the knitting process runs inside a temporary folder that is different from the current working directory.

# A dirty hack

The only way I've found to get `checkpoint()` to work inside an RMardown document, is really a bit of a hack. The workaround is to create a manifest of required packages, and write this to an R file in the working directory.

For example, imagine your R script uses the `MASS` package, then create a manifest file like this:

```{r}
td <- dirname(tempdir())
td <- getwd()
dir.create(file.path(td, ".checkpoint"), showWarnings = FALSE)
```


# Running checkpoint

Then you can use `checkpoint()` in the normal way.

In this example I use a temporary folder as the checkpoint location. This is only necessary because CRAN policy forbids writing anything to local disk, and a temporary folder is the only acceptable location.

```{r}
library(checkpoint)
checkpoint("2017-03-28", checkpointLocation = td)
```

# Confirm this actually worked:

If this worked, you should see that the library path now points to `tempdir()` and that MASS should be one of only a few package installed:

```{r}
.libPaths()
installed.packages(noCache = TRUE)[, "Package"]
```

# Your real R code:

Now your real R code follows

```{r}
library(MASS)
hist(islands)
```

# Conclusion

This is a bit of a dirty hack, but points in a direction for getting your RMarkdown script to be checkpointed.

Now your R Markdown script is checkpointed, and furthermore if you execute the R command `knitr::purl("foo.Rmd")` in a separate R session where `foo.Rmd` is the name of your R Markdown script, this will strip off only the R code associated with your project, but with the checkpoint facility intact, and write just the R code to a file named `foo.R`.

Now all you need to do is to circulate the file `foo.R` along with any supporting data files.

```{r, include=FALSE}
# clean up

detach("package:MASS", unload = TRUE)
unlink(file.path(td, ".checkpoint"), recursive = TRUE)
```